---
title: "ANOVA Model"
author: Jiongqi Zhao
output: ioslides_presentation
---

```{r,echo=FALSE,warning=FALSE,message=FALSE}
library(tidyverse)
library(lubridate)

covid <- read_csv("https://covid19.who.int/WHO-COVID-19-global-data.csv")
covid$WHO_region<-as.factor(covid$WHO_region)
```

```{r,echo=FALSE,message=FALSE,include=FALSE}
data3<-read_csv("C:/Users/liufe/Desktop/207/data_c.csv",col_names = T)
data3f<-read_csv("C:/Users/liufe/Desktop/207/data_c1.csv",col_names = T)
colnames(data3f)[1]<-"country_code"
colnames(data3)[1]<-"country_code"
sumn<-function(x){
  return (sum(x,na.rm=T))
}
data3f$days<-apply(data3f[,3:431],1,sumn)
pos<-function(x){
  return (which(x==2))
}
data3_pos<-apply(data3[,3:431],1,pos)
names(data3_pos)<-1:186

data3_pos1<-sapply(data3_pos,length)
data3_pos1<-data3_pos1[which(data3_pos1>=22)]
data3_pos<-data3[as.integer(names(data3_pos1)),]
posq<-function(x){
  return (which(x==2)[1])
}
pos1<-apply(data3_pos[,3:431],1,posq)
date<-colnames(data3_pos)[pos1+2]
Sys.setlocale("LC_TIME", "C")
date<-as.Date(date,"%d-%b-%y")
date<-data.frame(country_code=data3_pos$country_code,date=date)
library(countrycode)
library(ISOcodes)
covid_<-covid
covid_$country_code<-countrycode(covid$Country,origin = "country.name",destination = "iso3c",warn = TRUE)
c<-c(integer(0))
for (i in 1:153){
  c<-c(c,which(covid_$country_code==data3_pos$country_code[i]))
  
}

covid_<-covid_[c,]
newd<-inner_join(covid_, data3_pos, by = "country_code")
newd<-inner_join(newd,date, by = "country_code")
newd<-newd[,c(1:10,440)]

day<-newd[seq(1,64904,by=427),]$date
code<-unique(newd$country_code)
```
## Model
- growth rate = $\frac{New Cases in Day i}{Cumulative Cases in Day i-1}$
- $Y_{ijk} = \mu_{\cdot\cdot} + \alpha_i+\beta_j +\epsilon_{ijk}$,
where Y represents the growth rate of COVID-19, the number of observations in each cell is k = 1,..., 12, $\alpha$ denotes the lockdown status, i = 0 (no lockdown), 1 (lockdown), and $\beta$ denotes 3 country codes of Australia, United Kingdom, and India, we choose them because they have nationwise lockdowns. The constraints are (i)$\sum\alpha_i=0$, (ii)$\sum\beta_j=0$, (iii)$\epsilon_{ijk}$ follows $N(0,\sigma^2)$.
- The model assumptions are that the residuals follow a normal distribution and have equal variances.

## ANOVA
```{r,echo=FALSE,warning=FALSE}
data41<-data.frame(integer(0))
n=12
for (i in 1:152){
covid_temp<-filter(newd,country_code==code[i])

covid_temp_un<-filter(covid_temp,Date_reported<=date[i])
covid_temp_lo<-filter(covid_temp,date[i]<=Date_reported,Date_reported<=as.Date(date[i])+n)

covid_temp_un$rate<-c(NA,covid_temp_un$New_cases[2:length(covid_temp_un$New_cases)]/covid_temp_un$Cumulative_cases[1:length(covid_temp_un$New_cases)-1])
covid_temp_lo$rate<-c(NA,covid_temp_lo$New_cases[2:(n+1)]/covid_temp_lo$Cumulative_cases[1:n])
covid_temp_un$status<-rep(0,length(covid_temp_un$New_cases))
covid_temp_lo$status<-rep(1,(n+1))

temp1<-as.data.frame(rbind(covid_temp_un[2:length(covid_temp_un$New_cases),],covid_temp_lo[2:(n+1),]))
data41<-as.data.frame(rbind(data41,temp1))
}

#pos21<-which(is.nan(data41$rate)==|data41$rate==Inf)
#data41<-filter(data41,!(country_code%in%unique(data41$country_code[pos21])))

newdata<-filter(data41,(Country_code%in%c("US","UK","IN","AU","ET")))
UK<-filter(data41,(Country_code%in%c("GB")))
IN<-filter(data41,(Country_code%in%c("IN")))
AU<-filter(data41,(Country_code%in%c("AU")))
#ET<-filter(data41,(Country_code%in%c("ET")))
#US<-filter(data41,(Country_code%in%c("US")))

data<-UK[30:(length(UK$rate)-n),]$rate
resm4<-arima(data[30:(length(UK$rate)-n)], order=c(6,0,0))
pred4 <- predict(resm4, n.ahead=n, se.fit=TRUE)
fdat<-data.frame(new_cases=c(pred4$pred,UK[(length(UK$rate)-n+1):length(UK$rate),]$rate),status=c(rep(0,n),rep(1,n)),Country=rep("UK",(2*n)))

datain<-IN[30:(length(IN$rate)-n),]$rate
resm4<-arima(datain[30:(length(IN$rate)-n)], order=c(1,0,0))
pred4 <- predict(resm4, n.ahead=n, se.fit=TRUE)
fdatin<-data.frame(new_cases=c(pred4$pred,IN[(length(IN$rate)-n+1):length(IN$rate),]$rate),status=c(rep(0,n),rep(1,n)),Country=rep("IN",(2*n)))

dataau<-AU[24:(length(AU$rate)-n),]$rate
resm4<-arima(dataau[24:(length(AU$rate)-n)], order=c(4,0,0))
pred4 <- predict(resm4, n.ahead=n, se.fit=TRUE)
fdatau<-data.frame(new_cases=c(pred4$pred,AU[(length(AU$rate)-n+1):length(AU$rate),]$rate),status=c(rep(0,n),rep(1,n)),Country=rep("AU",(2*n)))

new<-as.data.frame(rbind(fdat,fdatin,fdatau))
new$Country<-as.factor(new$Country)
fit1<-aov(new_cases~Country+status+Country:status,data=new)
fit2<-aov(new_cases~Country+status,data=new)
#anova(fit1,fit2)
summary(fit1)

```


## Causal Inference

```{r,echo=F,fig.align='center'}
new$status<-as.factor(new$status)

new$status = relevel(new$status,ref='0')
fit4 = glm(status~Country,family = binomial,data = new)
prob = fit4$fitted.values
pscore = ifelse(new$new_cases=='1',prob,(1-prob))
weight = 1/pscore

model_new1= aov(new_cases~status+Country+status:Country,data = new,weights = weight)
summary(model_new1)
```

## Diagnostics

```{r,echo=FALSE}
par(mfrow=c(1,2))
plot(fit1, which=c(1,2))
```

## Conclusion

- We utilized the Logistic Growth Model to explore the trajectory of cumulative cases of COVID-19. 
- By the results of the AR and ANOVA models, there exists differences in the growth rate between different lockdown status and between different countries. 
