---
title: "Growth models"
author: "Team 4, Zhongxuan Liu"
date: "3/15/2021"
output: 
  ioslides_presentation:
        widescreen: yes
  beamer_presentation: default
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(warning=FALSE, message=FALSE, error = FALSE)
library(tidyverse)
library(ggplot2)
library(gridExtra)
library(FlexParamCurve)
covid <- read_csv("WHO-COVID-19-global-data.csv")
```

## Growth models

- Classic Logistic growing Model (CLM)

$$
    \dfrac{\mathrm{d}C(t)}{\mathrm{d}t} = rC(t)\left(1-\dfrac{C(t)}{K} \right) + \epsilon'(t)
$$

- Generalized Richards model (GRM)

$$
    \dfrac{\mathrm{d}C(t)}{\mathrm{d}t} = r[C(t)]^p \left(1-\left(\dfrac{C(t)}{K} \right)^\alpha \right) + \epsilon'(t)
$$

* $t$: time; $C(t)$: Cumulative cases; $r$: growth rate of cases
* $K$: carrying capacity; the upper asymptote of the pandemic
* $p\in [0,1]$ : capture the different growth profiles
* $\alpha$: capture the deviation from S-shaped curve


## Visulization of the growth models

```{r}
K <- 100000; r <- 0.2; alpha <- 0.8; p <- 1; C0 <- 10
demo1 <- function(t) K*C0*exp(r*t)/(K+C0*(exp(r*t)-1))
Q <- -1 + (K/C0)^alpha
demo2 <- function(t) K/(1+ Q*exp(-r*alpha*t))^(1/alpha)

t <- 1:100
demo_data <- data.frame(days = t)

ggplot(demo_data, aes(x = days)) +
    geom_function(aes(colour = "CLM"), size = 1.2, fun = demo1) + 
    geom_function(aes(colour = "GRM"), size = 1.2, fun = demo2) + 
    theme(legend.position = c(0.8, 0.2)) + 
    labs(x = "Day number", y = "Cumulative cases", 
        title = "Model trajectories (K = 1e5, alpha = 0.8, p = 1)")
```

## Modeling the COVID-19 infection trajectory

```{r include=FALSE}
trend_plot <- function(country) {
    clm <- nls(Cumulative_cases ~ SSlogis(Date_num, phi1, phi2, phi3), data = country)
    alpha <- coef(clm)
    modpar(country$Date_num, country$Cumulative_cases, verbose = T,
           force4par = T, suppress.text = T, pn.options = "myoptions", width.bounds = 100)
    grm <- nls(Cumulative_cases ~ SSposnegRichards(Date_num, Asym = Asym, K = K, 
                                                   Infl = Infl, M = M, 
                                                   pn.options = "myoptions", modno = 12),
               data = country)
    beta <- coef(grm)
    
    f1 <- function(x) alpha[1]/(1 + exp(-(x - alpha[2])/alpha[3]))
    f2 <- function(x) posnegRichards.eqn(x, Asym = beta[1], K = beta[2], Infl = beta[3], 
                                         M = beta[4], pn.options = "myoptions", modno = 12)
    if(length(country) == 6){
        country_name <- "world"
    }
    else{
        country_name <- country$Country[1]
    }
    
    plot_country <- ggplot(country, aes(x = Date_num, y = Cumulative_cases, colour = "Real")) +
        geom_point() + geom_line() +
        geom_function(aes(colour = "CLM"), size = 1.2, fun = f1) + 
        geom_function(aes(colour = "GRM"), size = 1.2, fun = f2) + 
        theme(legend.position = c(0.1, 0.8)) + 
        labs(x = "Day number", y = "Cumulative cases", 
             title = paste("COVID trend in", country_name))
    
    return(plot_country)
}

covid_global <- covid %>% group_by(Date_reported) %>%
    summarise(across(where(is.numeric), sum))
covid_global <- covid_global %>% mutate(
    Date_num = as.numeric(difftime(Date_reported, Date_reported[1], units = "days")))
covid_global <- head(covid_global, -1)
global_plot <- trend_plot(covid_global)
global_plot$labels$title <- "COVID trend in the world"

covid_us <- covid %>% filter(Country_code == "US")
covid_us <- covid_us %>% mutate(
    Date_num = as.numeric(difftime(Date_reported, Date_reported[1], units = "days")))
us_plot <- trend_plot(covid_us)

covid_gb <- covid %>% filter(Country_code == "GB")
covid_gb <- covid_gb %>% mutate(
    Date_num = as.numeric(difftime(Date_reported, Date_reported[1], units = "days")))
gb_plot <- trend_plot(covid_gb)

covid_in <- covid %>% filter(Country == "India")
covid_in <- covid_in %>% mutate(
    Date_num = as.numeric(difftime(Date_reported, Date_reported[1], units = "days")))
in_plot <- trend_plot(covid_in)

covid_et <- covid %>% filter(Country == "Ethiopia")
covid_et <- covid_et %>% mutate(
    Date_num = as.numeric(difftime(Date_reported, Date_reported[1], units = "days")))
covid_et <- head(covid_et, -1)
et_plot <- trend_plot(covid_et)

covid_au <- covid %>% filter(Country == "Australia")
covid_au <- covid_au %>% mutate(
    Date_num = as.numeric(difftime(Date_reported, Date_reported[1], units = "days")))
covid_au <- head(covid_au, -1)
au_plot <- trend_plot(covid_au)

fig <- grid.arrange(global_plot, us_plot, gb_plot, in_plot, et_plot, au_plot,
             nrow = 2)
```


```{r trendplot, fig.height=5, fig.width=10}
grid.arrange(fig)
```


## Diagnostics for the growth models

The models failed to converge in some countries: sensitive to the data.

```{r, fig.height=5, fig.width=10}
covid_cn <- covid %>% filter(Country_code == "CN")
covid_cn <- covid_cn %>% mutate(
    Date_num = as.numeric(difftime(Date_reported, Date_reported[1], units = "days")))
cn_plot <- ggplot(covid_cn, aes(x = Date_num, y = Cumulative_cases)) + 
    geom_point() + geom_line() + 
    labs(x = "Day number", y = "Cumulative cases", 
        title = "Trend in China")

covid_za <- covid %>% filter(Country == "South Africa")
covid_za <- covid_za %>% mutate(
    Date_num = as.numeric(difftime(Date_reported, Date_reported[1], units = "days")))
za_plot <- ggplot(covid_za, aes(x = Date_num, y = Cumulative_cases)) + 
    geom_point() + geom_line() + 
    labs(x = "Day number", y = "Cumulative cases", 
        title = "Trend in South Africa")

covid_br <- covid %>% filter(Country == "Brazil")
covid_br <- covid_br %>% mutate(
    Date_num = as.numeric(difftime(Date_reported, Date_reported[1], units = "days")))
br_plot <- ggplot(covid_br, aes(x = Date_num, y = Cumulative_cases)) + 
    geom_point() + geom_line() + 
    labs(x = "Day number", y = "Cumulative cases", 
        title = "Trend in Brazil")

grid.arrange(cn_plot, za_plot, br_plot, 
             nrow = 1)
```

## Residual plot for the CLM model and the global cases

The residuals of the growth model is relatively big

```{r}
w_clm <- nls(Cumulative_cases ~ SSlogis(Date_num, phi1, phi2, phi3), data = covid_global)
w_alpha <- coef(w_clm)
fw <- function(x) w_alpha[1]/(1 + exp(-(x - w_alpha[2])/w_alpha[3]))
res_data <- covid_global$Date_num
res <- fw(covid_global$Date_num) - covid_global$Cumulative_cases

ggplot(covid_global, aes(x = Cumulative_cases, y = res)) + geom_point()
```

## 

- It should be used to model the long-term trajectory of COVID-19.

- It might not be appropriate to use growth models to predict short-term cases. 

- To research on the short-term influence of policies such as lockdown, we use ANOVA and AR model.