---
title: "Statistical Data Analysis On WHO Covid-19 data"
author: "Team 4: Jiawen Liu, Jiongqi Zhao, Zhongxuan Liu, Chenze Li"
date: "3/14/2021"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

# Descriptive Analysis

## Interactive plot of new cases by time

```{r, message=FALSE}
library(tidyverse)
library(plotly)
library(gplots)
library(lme4)
library(data.table)
library(kableExtra)
library(lubridate)
library(hrbrthemes)
covid <- read_csv("~/Desktop/Courses/UC Davis/2021winter/207/WHO-COVID-19-global-data.csv")
```

```{r, fig.align='center', message = FALSE}
# summarise by date
covid_date <- covid %>% 
  mutate(avg_new_cases = stats::filter(New_cases, rep(1/7, 7), sides = 1),
         avg_new_deaths = stats::filter(New_deaths, rep(1/7, 7), sides = 1)) %>% 
  group_by(Date_reported) %>% 
  summarise(sum_new_cases = sum(New_cases),
            sum_new_deaths = sum(New_deaths),
            sum_avg_new_cases = sum(avg_new_cases),
            sum_avg_new_deaths = sum(avg_new_deaths))
p1 <- plot_ly(data = covid_date, x = ~ Date_reported, y = ~ sum_new_cases, type = "bar", name = "new cases")
add_paths(p1, x = covid_date$Date_reported, y = covid_date$sum_avg_new_cases, name = "7 day average")
```


## Interactive plot of new deaths by time

```{r}
p2 <- plot_ly(data = covid_date, x = ~ Date_reported, y = ~ sum_new_deaths, type = "bar", name = "new deaths")
add_paths(p2, x = covid_date$Date_reported, y = covid_date$sum_avg_new_deaths, name = "7 day average")
```

## The six countries with the most cumulative cases so far

| Country   | Cumulative cases | Cumulative deaths |
| ----------- | ----------- | ----------- |
| United States of America  | 29063401   |      528456|
| Brazil  | 11363380    |     275105|
| India  | 11359048    |     158607|
| Russian Federation  | 4390608     |     92090|
| The United Kingdom | 4253824     |    125464|
| France  | 3975989    |      89800|

```{r, fig.align='center', message = FALSE}
covid_country <- covid %>% 
  group_by(Country) %>% 
  summarise(sum_new_cases = sum(New_cases),
            sum_new_deaths = sum(New_deaths)) %>% 
  arrange(desc(sum_new_cases))

# kable(head(covid_country))
```

## WHO region defination

| WHO_region   | Whole name |
| ----------- | ----------- |
| AFRO  | WHO Regional Office for Africa |
| AMRO  | WHO Regional Office for the Americas |
| EMRO  | WHO Regional Office for the Eastern Mediterranean |
| EURO  | WHO Regional Office for Europe |
| SEARO | WHO Regional Office for South-East Asia |
| WPRO  | WHO Regional Office for the Western Pacific |

Reference: https://www.who.int/chp/about/regions/en/

## Bar chart of cumulative cases by WHO region

```{r, message = FALSE}
covid_region <- covid %>% 
  group_by(WHO_region) %>% 
  summarise(sum_new_cases = sum(New_cases),
            sum_new_deaths = sum(New_deaths))

ggplot(covid_region, aes(x = factor(WHO_region, levels = c("Other", "AFRO", "AMRO", "EMRO",  "EURO", "SEARO", "WPRO")),y=sum_new_cases, fill = WHO_region)) + 
  geom_bar(stat="identity") + coord_flip() + xlab("WHO_region")
```

## The world map

```{r, fig.align='center', message=FALSE, fig.height=4, warning=FALSE}
library(tidyverse)
covid$WHO_region<-as.factor(covid$WHO_region)
#region<-group_by(covid,WHO_region,Date_reported)
data<-filter(covid,Date_reported=="2021-02-26")
region<-group_by(data,WHO_region)
datar<-summarise(region,mean_new=mean(New_cases),mean_cu=mean(Cumulative_cases),mean_de=mean(New_deaths),mean_desum=mean(Cumulative_deaths))

library(ggplot2)
library(dplyr)
library(countrycode)
library(ISOcodes)
require(maps)
require(viridis)
theme_set(
  theme_void()
)
#interactive plot
library(echarts4r)
# covid_es<-filter(covid,Date_reported>= "2021-01-03", Date_reported<= "2021-02-28")
covid_es <- covid
covid_es$isoa2<-countrycode(covid_es$Country,origin = "country.name",destination = "country.name",warn = TRUE)
covid_es <- covid_es %>%
mutate(Name = recode_factor(isoa2,
                            `Congo - Kinshasa`= "Dem. Rep. Congo",
                            `Congo - Brazzaville`= "Congo",
                            `Cote d'Ivoire`= "C??te d'Ivoire",
                            `Central African Republic`= "Central African Rep.",
                            `Yemen, Rep.`= "Yemen",
                            `Korea, Rep.`= "Korea",
                            `Korea, Dem. Rep.`= "Dem. Rep. Korea",
                            `Czech Republic`= "Czech Rep.",
                            `Slovak Republic`= "Slovakia",
                            `Dominican Republic`= "Dominican Rep.",
                            `Equatorial Guinea`= "Eq. Guinea",
                            `South Sudan`="S.Sudan",
                            `Myanmar (Burma)`="Myanmar"))
 covid_es%>%
  group_by(Date_reported) %>%
  e_chart(Name, timeline = TRUE) %>%
  e_map(New_cases) %>%
  e_visual_map(min=0,max=170000,type = 'continuous') %>%
  e_title("New Cases of Covid-19 by Country and Day", left = "center") %>%
  e_tooltip(
    trigger = "item",
    formatter = e_tooltip_choro_formatter())
```

```{r, fig.align='center', message=FALSE}
# # 
# covid_es <- covid
# covid_es$isoa2<-countrycode(covid_es$Country,origin = "country.name",destination = "country.name",warn = TRUE)
# covid_es <- covid_es %>%
# mutate(Name = recode_factor(isoa2,
#                             `Congo - Kinshasa`= "Dem. Rep. Congo",
#                             `Congo - Brazzaville`= "Congo",
#                             `Cote d'Ivoire`= "C??te d'Ivoire",
#                             `Central African Republic`= "Central African Rep.",
#                             `Yemen, Rep.`= "Yemen",
#                             `Korea, Rep.`= "Korea",
#                             `Korea, Dem. Rep.`= "Dem. Rep. Korea",
#                             `Czech Republic`= "Czech Rep.",
#                             `Slovak Republic`= "Slovakia",
#                             `Dominican Republic`= "Dominican Rep.",
#                             `Equatorial Guinea`= "Eq. Guinea",
#                             `South Sudan`="S.Sudan",
#                             `Myanmar (Burma)`="Myanmar")) %>% 
#   filter(Country %in% c("United States of America", "The United Kingdom", "India", "Australia", "Ethiopia"))
#  covid_es%>%
#   group_by(Date_reported) %>%
#   e_chart(Name, timeline = TRUE) %>%
#   e_map(New_cases) %>%
#   e_visual_map(min=0,max=170000,type = 'continuous') %>%
#   e_title("New Cases of Covid-19 by Country and Day", left = "center") %>%
#   e_tooltip(
#     trigger = "item",
#     formatter = e_tooltip_choro_formatter())
```

## Interactive plot of new cases by time of USA, UK

```{r, fig.height=2, message=FALSE}
# America
covid <- read_csv("~/Desktop/Courses/UC Davis/2021winter/207/WHO-COVID-19-global-data.csv")
covid_us <- covid %>% 
  mutate(avg_new_cases = stats::filter(New_cases, rep(1/7, 7), sides = 1),
         avg_new_deaths = stats::filter(New_deaths, rep(1/7, 7), sides = 1)) %>% 
  filter(Country == "United States of America")

p3 <- plot_ly(data = covid_us, x = ~ Date_reported, y = ~ New_cases, type = "bar") %>% layout(title = "US")
p3

# The United Kingdom
covid_uk <- covid %>% 
  mutate(avg_new_cases = stats::filter(New_cases, rep(1/7, 7), sides = 1),
         avg_new_deaths = stats::filter(New_deaths, rep(1/7, 7), sides = 1)) %>% 
  filter(Country == "The United Kingdom")

p4 <- plot_ly(data = covid_uk, x = ~ Date_reported, y = ~ New_cases, type = "bar") %>% layout(title = "UK")
p4
```

## Interactive plot of new cases by time of India, Australia
```{r, fig.height=2}
# India
covid_in <- covid %>% 
  mutate(avg_new_cases = stats::filter(New_cases, rep(1/7, 7), sides = 1),
         avg_new_deaths = stats::filter(New_deaths, rep(1/7, 7), sides = 1)) %>% 
  filter(Country == "India")

p5 <- plot_ly(data = covid_in, x = ~ Date_reported, y = ~ New_cases, type = "bar") %>% layout(title = "India")
p5

# Australia
covid_aus <- covid %>% 
  mutate(avg_new_cases = stats::filter(New_cases, rep(1/7, 7), sides = 1),
         avg_new_deaths = stats::filter(New_deaths, rep(1/7, 7), sides = 1)) %>% 
  filter(Country == "Australia")

p6 <- plot_ly(data = covid_aus, x = ~ Date_reported, y = ~ New_cases, type = "bar") %>% layout(title = "Australia")
p6
```

## Interactive plot of new cases by time of Ethiopia

```{r, fig.height=2}
# Ethiopia
covid_eth <- covid %>% 
  mutate(avg_new_cases = stats::filter(New_cases, rep(1/7, 7), sides = 1),
         avg_new_deaths = stats::filter(New_deaths, rep(1/7, 7), sides = 1)) %>% 
  filter(Country == "Ethiopia")

p7 <- plot_ly(data = covid_eth, x = ~ Date_reported, y = ~ New_cases, type = "bar") %>% layout(title = "Ethiopia")
p7
```

## Growth rate by month
```{r, message=FALSE}
covid.growth <- covid %>% 
  mutate(year = year(Date_reported),
         month = month(Date_reported)) %>% 
  group_by(year, month) %>% 
  summarise(sum_new_cases = sum(New_cases)) %>% 
  ungroup()

covid.growth.new <- covid.growth %>% 
  mutate(temp = lag(sum_new_cases, n = 1),
         Time = paste(year, month, sep = "-"),
         Time = paste0(Time, "-01"),
         #temp = ifelse(Time == "2021-1", 19082911, temp),
         growth_rate = (sum_new_cases-temp)/sum_new_cases,
         growth_rate = ifelse(is.na(growth_rate), 0, growth_rate))

covid.growth.new$Month <- ymd(covid.growth.new$Time)
covid.growth.new <- covid.growth.new %>% 
  filter(Month != "2021-02-01") %>% 
  filter(Month != "2021-03-01")
covid.growth.new <- data.frame(covid.growth.new)
ggplot(covid.growth.new, aes(x = Month, y = growth_rate)) + 
  geom_line() + xlab("Time") + theme_ipsum() + ggtitle("Monthly growth rate")
```

一

