---
title: "Project3-Presentation-Team4"
author: "Jiawen Liu, Jiongqi Zhao, Zhongxuan Liu, Chenze Li"
date: "3/14/2021"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

# Descriptive Analysis

## Interactive plot of new cases by time

```{r, message=FALSE}
library(tidyverse)
library(plotly)
library(gplots)
library(lme4)
library(data.table)
covid <- read_csv("~/Desktop/Courses/UC Davis/2021winter/207/WHO-COVID-19-global-data.csv")
```

```{r, fig.align='center', message = FALSE}
# summarise by date
covid_date <- covid %>% 
  mutate(avg_new_cases = stats::filter(New_cases, rep(1/7, 7), sides = 1),
         avg_new_deaths = stats::filter(New_deaths, rep(1/7, 7), sides = 1)) %>% 
  group_by(Date_reported) %>% 
  summarise(sum_new_cases = sum(New_cases),
            sum_new_deaths = sum(New_deaths),
            sum_avg_new_cases = sum(avg_new_cases),
            sum_avg_new_deaths = sum(avg_new_deaths))
p1 <- plot_ly(data = covid_date, x = ~ Date_reported, y = ~ sum_new_cases, type = "bar", name = "new cases")
add_paths(p1, x = covid_date$Date_reported, y = covid_date$sum_avg_new_cases, name = "7 day average")
```


## Interactive plot of new deaths by time

```{r}
p2 <- plot_ly(data = covid_date, x = ~ Date_reported, y = ~ sum_new_deaths, type = "bar", name = "new deaths")
add_paths(p2, x = covid_date$Date_reported, y = covid_date$sum_avg_new_deaths, name = "7 day average")
```

## The six countries with the most cumulative cases so far

```{r, fig.align='center', message = FALSE}
covid_country <- covid %>% 
  group_by(Country_code) %>% 
  summarise(sum_new_cases = sum(New_cases),
            sum_new_deaths = sum(New_deaths)) %>% 
  arrange(desc(sum_new_cases))

head(covid_country)
```

## Bar chart of cumulative cases by WHO region

```{r, message = FALSE}
covid_region <- covid %>% 
  group_by(WHO_region) %>% 
  summarise(sum_new_cases = sum(New_cases),
            sum_new_deaths = sum(New_deaths))

ggplot(covid_region, aes(x = factor(WHO_region, levels = c("Other", "AFRO", "AMRO", "EMRO",  "EURO", "SEARO", "WPRO")),y=sum_new_cases, fill = WHO_region)) + 
  geom_bar(stat="identity") + coord_flip()
```

## Interactive plot of new cases by time of USA, UK

```{r, fig.height=2}
# America
covid_us <- covid %>% 
  mutate(avg_new_cases = stats::filter(New_cases, rep(1/7, 7), sides = 1),
         avg_new_deaths = stats::filter(New_deaths, rep(1/7, 7), sides = 1)) %>% 
  filter(Country == "United States of America")

p3 <- plot_ly(data = covid_us, x = ~ Date_reported, y = ~ New_cases, type = "bar") %>% layout(title = "US")
p3

# The United Kingdom
covid_uk <- covid %>% 
  mutate(avg_new_cases = stats::filter(New_cases, rep(1/7, 7), sides = 1),
         avg_new_deaths = stats::filter(New_deaths, rep(1/7, 7), sides = 1)) %>% 
  filter(Country == "The United Kingdom")

p4 <- plot_ly(data = covid_uk, x = ~ Date_reported, y = ~ New_cases, type = "bar") %>% layout(title = "UK")
p4
```

## Interactive plot of new cases by time of India, Australia and Ethiopia
```{r, fig.height=2}
# India
covid_in <- covid %>% 
  mutate(avg_new_cases = stats::filter(New_cases, rep(1/7, 7), sides = 1),
         avg_new_deaths = stats::filter(New_deaths, rep(1/7, 7), sides = 1)) %>% 
  filter(Country == "India")

p5 <- plot_ly(data = covid_in, x = ~ Date_reported, y = ~ New_cases, type = "bar") %>% layout(title = "India")
p5

# Australia
covid_aus <- covid %>% 
  mutate(avg_new_cases = stats::filter(New_cases, rep(1/7, 7), sides = 1),
         avg_new_deaths = stats::filter(New_deaths, rep(1/7, 7), sides = 1)) %>% 
  filter(Country == "Australia")

p6 <- plot_ly(data = covid_aus, x = ~ Date_reported, y = ~ New_cases, type = "bar") %>% layout(title = "Australia")
p6
```

## Interactive plot of new cases by time of Ethiopia

```{r, fig.height=2}
# Ethiopia
covid_eth <- covid %>% 
  mutate(avg_new_cases = stats::filter(New_cases, rep(1/7, 7), sides = 1),
         avg_new_deaths = stats::filter(New_deaths, rep(1/7, 7), sides = 1)) %>% 
  filter(Country == "Ethiopia")

p7 <- plot_ly(data = covid_eth, x = ~ Date_reported, y = ~ New_cases, type = "bar") %>% layout(title = "Ethiopia")
p7
```

## The world map

```{r, fig.align='center', message=FALSE}
library(tidyverse)
covid <- read_csv("https://covid19.who.int/WHO-COVID-19-global-data.csv")
covid$WHO_region<-as.factor(covid$WHO_region)
#region<-group_by(covid,WHO_region,Date_reported)
data<-filter(covid,Date_reported=="2021-02-26")
region<-group_by(data,WHO_region)
datar<-summarise(region,mean_new=mean(New_cases),mean_cu=mean(Cumulative_cases),mean_de=mean(New_deaths),mean_desum=mean(Cumulative_deaths))

library(ggplot2)
library(dplyr)
library(countrycode)
library(ISOcodes)
require(maps)
require(viridis)
theme_set(
  theme_void()
)
#interactive plot
library(echarts4r)
# covid_es<-filter(covid,Date_reported>= "2021-01-03", Date_reported<= "2021-02-28")
covid_es <- covid
covid_es$isoa2<-countrycode(covid_es$Country,origin = "country.name",destination = "country.name",warn = TRUE)
covid_es <- covid_es %>%
mutate(Name = recode_factor(isoa2,
                            `Congo - Kinshasa`= "Dem. Rep. Congo",
                            `Congo - Brazzaville`= "Congo",
                            `Cote d'Ivoire`= "C??te d'Ivoire",
                            `Central African Republic`= "Central African Rep.",
                            `Yemen, Rep.`= "Yemen",
                            `Korea, Rep.`= "Korea",
                            `Korea, Dem. Rep.`= "Dem. Rep. Korea",
                            `Czech Republic`= "Czech Rep.",
                            `Slovak Republic`= "Slovakia",
                            `Dominican Republic`= "Dominican Rep.",
                            `Equatorial Guinea`= "Eq. Guinea",
                            `South Sudan`="S.Sudan",
                            `Myanmar (Burma)`="Myanmar"))
 covid_es%>%
  group_by(Date_reported) %>%
  e_chart(Name, timeline = TRUE) %>%
  e_map(New_cases) %>%
  e_visual_map(min=0,max=170000,type = 'continuous') %>%
  e_title("New Cases of Covid-19 by Country and Day", left = "center") %>%
  e_tooltip(
    trigger = "item",
    formatter = e_tooltip_choro_formatter())
```

## 不知道叫啥

```{r, fig.align='center', message=FALSE}
# 
covid_es <- covid
covid_es$isoa2<-countrycode(covid_es$Country,origin = "country.name",destination = "country.name",warn = TRUE)
covid_es <- covid_es %>%
mutate(Name = recode_factor(isoa2,
                            `Congo - Kinshasa`= "Dem. Rep. Congo",
                            `Congo - Brazzaville`= "Congo",
                            `Cote d'Ivoire`= "C??te d'Ivoire",
                            `Central African Republic`= "Central African Rep.",
                            `Yemen, Rep.`= "Yemen",
                            `Korea, Rep.`= "Korea",
                            `Korea, Dem. Rep.`= "Dem. Rep. Korea",
                            `Czech Republic`= "Czech Rep.",
                            `Slovak Republic`= "Slovakia",
                            `Dominican Republic`= "Dominican Rep.",
                            `Equatorial Guinea`= "Eq. Guinea",
                            `South Sudan`="S.Sudan",
                            `Myanmar (Burma)`="Myanmar")) %>% 
  filter(Country %in% c("United States of America", "The United Kingdom", "India", "Australia", "Ethiopia"))
 covid_es%>%
  group_by(Date_reported) %>%
  e_chart(Name, timeline = TRUE) %>%
  e_map(New_cases) %>%
  e_visual_map(min=0,max=170000,type = 'continuous') %>%
  e_title("New Cases of Covid-19 by Country and Day", left = "center") %>%
  e_tooltip(
    trigger = "item",
    formatter = e_tooltip_choro_formatter())
```





